---
title: "Paper_figures_Axin2_suture"
subtitle: "Fig.1B"
author: "Zhirui Jiang"
date: "2023-6-22"
output: 
  html_document: 
    fig_width: 12
    fig_height: 9
editor_options: 
  chunk_output_type: console
---
```{r}
library(Seurat)
library(tidyverse)
library(rJava)
devtools::install_github('colearendt/xlsx')
library(xlsx)
library(ggplot2)
```

# Fig. 1B.Pathway enrichment analysis of differentially expressed genes (DEGs) in SC cluster exhibited by Circular Histogram showing the top 20 enriched KEGG pathways of genes expressed and by Bar graph showing selected pathways known related to stem cell regulation from the top 100 enriched KEGG pathways.  

# Draw circle plot for KEGG pathway enrichment analysis
```{r}
testseu = readRDS("Axin2+_scRNA-seq_combined.rds")
# Idents(testseu) <- "anno_new"
```

```{r}
### Find DEGs 
marker_celltype <- FindAllMarkers(
  testseu,
  only.pos = TRUE,
  min.pct =  0.25,
  min.diff.pct = 0.25
)

# filter
marker_celltype = marker_celltype %>% filter(p_val_adj < 0.05)
marker_celltype$d = marker_celltype$pct.1 - marker_celltype$pct.2
marker_celltype = marker_celltype %>% filter(d > 0.2)
marker_celltype = marker_celltype %>% arrange(cluster, desc(avg_log2FC))
marker_celltype = as.data.frame(marker_celltype)
write.xlsx(
  marker_celltype,
  file = "markers_log2fc0.25_padj0.05_pctd0.2.xlsx",
  row.names = F,
  col.names = T
)
```

```{r}
### enrichment analysis
library(clusterProfiler)
library(org.Mm.eg.db)
R.utils::setOption("clusterProfiler.download.method", "auto") #https://github.com/YuLab-SMU/clusterProfiler/issues/256

source("syEnrich.R")
syEnrich(marker_celltype, outpath = "markers_log2fc0.25_padj0.05_pctd0.2")

```

```{r}
### find one cell type
kegg.res = read.xlsx(
  "markers_log2fc0.25_padj0.05_pctd0.2.KEGG.xls",
  sheetIndex = 1,
  as.data.frame = T,
  header = T
)
kegg.res = kegg.res %>% filter(p.adjust < 0.05)
kegg.res = kegg.res %>% filter(cluster == "SC")

# import kegg_info
kegg_info = read.xlsx("kegg_info.xlsx", sheetIndex = 1, startRow = 3)
kegg_info = kegg_info[, c("ID", "Pathway", "big.annotion")]

# merge
kegg.res$ID = str_replace(kegg.res$ID, "mmu", "")
kegg.res = kegg.res %>% inner_join(kegg_info, by = "ID")

# top20 term
kegg.res = kegg.res %>% arrange(p.adjust)
kegg.res = head(kegg.res, 20)

saveRDS(kegg.res, file = "kegg.res.rds")

write.table(
  kegg.res,
  file = "kegg.res.txt",
  quote = F,
  sep = "\t",
  row.names = F,
  col.names = T
)
write.xlsx(kegg.res,
           file = "kegg.res.xlsx",
           col.names = T,
           row.names = F)

```

```{r}
### use kegg_loop as function
source("kegg_loop.R")
kegg_loop(enrich.res = kegg.res, filename = "SC_kegg_id")
```


## Bar plot showing selected KEGG or Reactome pathways among top 100
# Submit DEGs to DAVID https://davidbioinformatics.nih.gov/tools.jsp and download KEGG and Reactome results as .txt
```{r}
# KEGG/GO

dat = read.table("cluster4_KEGG.txt", sep='\t', header=T)

#order in counts
dat = dat[order(dat$Count,decreasing=T),]
nrows = nrow(dat)

#find p<0.05 GO term
dat = dat[which(dat$PValue < 0.05),]

#show top 20
show = dat[1:20,]
show = dat[c(4,14,15,18,21,24,45,62),]

#draw plot
gg = ggplot(show, aes(x=Term,y=Count,fill=PValue)) + geom_bar(width=0.8,stat="identity") + coord_flip() + scale_fill_gradient(low = "red", high = "green") + scale_x_discrete(limits=show$Term[8:1])
pdf("barplot_KEGG.pdf", width=12, height=6)
gg + theme(axis.text.y = element_text(size=20),axis.title=element_text(size=20),legend.text=element_text(size=20))
dev.off()

```

```{r}
# Reactome

dat = read.table("cluster4_Reactome.txt", sep='\t', header=T)

#order in counts
dat = dat[order(dat$Count,decreasing=T),]
nrows = nrow(dat)

#find p<0.05 GO term
dat = dat[which(dat$PValue < 0.05),]

#show top 20
show = dat[1:20,]
show = dat[c(4,7,14,25,26,39,43,48,51,67,82,83,84,85),]

#draw plot
gg = ggplot(show, aes(x=Term,y=Count,fill=PValue)) + geom_bar(width=0.8,stat="identity") + coord_flip() + scale_fill_gradient(low = "red", high = "green") + scale_x_discrete(limits=show$Term[14:1])
pdf("barplot_Reactome.pdf", width=12, height=6)
gg + theme(axis.text.y = element_text(size=20),axis.title=element_text(size=20),legend.text=element_text(size=20))
dev.off()
```


