---
title: "Paper_figures_Axin2_suture"
subtitle: "Fig.1-3, Fig.S2,S3,S5-7"
author: "Zhirui Jiang"
date: '2023-06-22'
output: 
  html_document: 
    fig_width: 12
    fig_height: 9
editor_options: 
  chunk_output_type: console
---
# Load library
```{r}
library(dplyr)
library(ggplot2)
library(Seurat)
library(tidyverse)
library(data.table)
library(magrittr)
library(patchwork)
library(ComplexHeatmap)
set.seed(1234)
```

# load data and filtering 
```{r}
Axin2_scRNAseq.data1<- Read10X(data.dir = "scRNA-seq Axin2 positive cell suture-first/outs/filtered_feature_bc_matrix/")
Axin2_scRNAseq.data2<- Read10X(data.dir = "scRNA-seq Axin2 positive cell suture-second/outs/filtered_feature_bc_matrix/")

Axin2.batch1<-CreateSeuratObject(counts = Axin2_scRNAseq.data1, project = "Axin2.batch1", min.cells = 3, min.features = 200)
Axin2.batch2<-CreateSeuratObject(counts = Axin2_scRNAseq.data2, project = "Axin2.batch2", min.cells = 3, min.features = 200)


head(x=Axin2.batch1@meta.data,5)
head(x=Axin2.batch2@meta.data,5)


suture.list<-list(Axin2.batch1,Axin2.batch2)
names(suture.list)<-c("batch1","batch2")

#Normalization of the data
#Identify variable features
suture.list<- lapply(X= suture.list, FUN=function(x){
x<-NormalizeData(object = x, normalization.method = "LogNormalize", scale.factor = 1e4)
x<-FindVariableFeatures(x,selection.method = "vst", nfeatures = 2000)
})

##Perform integration
Axin2.anchors<-FindIntegrationAnchors(object.list = suture.list,dims = 1:20)
Axin2.combined <- IntegrateData(anchorset = Axin2.anchors, dims = 1:20)

# Filter out low quality reads using selected thresholds - these will change with experiment
Axin2.data$log10GenesPerUMI <- log10(Axin2.data$nFeature_RNA) / log10(Axin2.data$nCount_RNA)
Axin2.data <- PercentageFeatureSet(Axin2.data, pattern = "^mt-", col.name = "percent.mt")
Axin2.data <- subset(x = Axin2.data, 
                         subset= (nCount_RNA >= 500) & 
                           (nFeature_RNA >= 500) & 
                           (log10GenesPerUMI > 0.76) & 
                           (percent.mt <= 15))

#Visualize QC Metrics as a violin plot
VlnPlot(object = Axin2.combined, features = c("nFeature_RNA", "nCount_RNA"), ncol=2, pt.size = 0)
plot1<- FeatureScatter(object = Axin2.combined, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1

##Perform an integrated analysis
DefaultAssay(Axin2.combined)<-"integrated"

#Scaling the data
#Shifts the expression of each gene, so that the mean expression across cells is 0 Scales the expression of each gene, so that the variance across cells is 1 The results of this are stored in pbmc[[“RNA”]]@scale.data
all.genes<-rownames(x=Axin2.combined)
class(all.genes)
head(all.genes)
Axin2.combined <- ScaleData(object = Axin2.combined, features = all.genes, verbose = FALSE)
Axin2.combined <- FindVariableFeatures(object = Axin2.combined)
Axin2.combined <- RunPCA(object = Axin2.combined, upcs = 30, features = VariableFeatures(object=Axin2.combined), verbose = FALSE)

# UMAP
Axin2.combined <- RunUMAP(Axin2.combined, reduction = "pca", dims = 1:20)
DimPlot(Axin2.combined,reduction="umap",split.by = "orig.ident", pt.size = 0.5)
VlnPlot(object = Axin2.combined, features = c("nFeature_RNA", "nCount_RNA"), ncol=2, pt.size = 0)

saveRDS(Axin2.combined, file = "Axin2_combined_pre_clustering.rds")

# Clustering
Axin2.combined <- FindNeighbors(Axin2.combined, reduction = "pca", dims = 1:20)
Axin2.combined <- FindClusters(Axin2.combined, resolution = 0.1)

saveRDS(Axin2.combined, file = "Axin2_combined.rds")
```

# Fig. S2. Quality control of scRNA-seq data for Axin2-expressing suture cells.  
```{r}
# Fig. S2A
VlnPlot(object = Axin2.combined, features = c("nFeature_RNA", "nCount_RNA","percent.mt"), ncol=3, pt.size = 0)
plot1<- FeatureScatter(object = Axin2.combined, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1

##Determine the dimentionality of the dataset
ElbowPlot(object = Axin2.combined)

all.genes<-rownames(x=Axin2.combined)
class(all.genes)
head(all.genes)
Axin2.combined <- ScaleData(object = Axin2.combined, features = all.genes, verbose = FALSE)
Axin2.combined <- FindVariableFeatures(object = Axin2.combined)
Axin2.combined <- RunPCA(object = Axin2.combined, upcs = 30, features = VariableFeatures(object=Axin2.combined), verbose = FALSE)
DefaultAssay(Axin2.combined)<-"integrated"
Axin2.combined <- RunUMAP(Axin2.combined, reduction = "pca", dims = 1:20)
Axin2.combined <- FindNeighbors(Axin2.combined, reduction = "pca", dims = 1:20)
Axin2.combined <- FindClusters(Axin2.combined, resolution = 0.1)


DimPlot(Axin2.combined,reduction = "umap",group.by = "orig.ident", pt.size = 0.6)

# Fig. S2B
DefaultAssay(Axin2.combined) <- "RNA"
all_markers <-FindAllMarkers(Axin2.combined, 
                             min.pct =  0.25, 
                             min.diff.pct = 0.25)
write.csv(all_markers, "all_markers_genes_res0.1.csv", quote = F)
all.genes <- rownames(Axin2.combined)
Axin2.combined <- ScaleData(Axin2.combined, features = all.genes)
DimHeatmap(Axin2.combined, dims = 1:3, cells = 500, balanced = TRUE)
all_markers %>%  
  group_by(cluster) %>% 
  top_n(n = 20, wt = avg_log2FC) -> top20
pdf("heatmap_20.pdf", width=6, height=12)
print(DoHeatmap(Axin2.combined, features = top20$gene) + NoLegend())
dev.off()
```

# Fig.1.Single-cell transcriptomic profiling of Axin2-expressing suture cells. 
```{r}
#Fig.1A
DimPlot(Axin2.combined,reduction="umap",split.by = "orig.ident")
new.cluster.ids<- c("NEUT1","NEUT2","MP","MONO","SC","MC")
names(x=new.cluster.ids) <-levels(x=Axin2.combined)
Axin2.combined<-RenameIdents(object = Axin2.combined,new.cluster.ids)
DimPlot(object = Axin2.combined,reduction = 'umap',label = TRUE,label.size = 4,pt.size = 0.5)+  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=16))

DefaultAssay(Axin2.combined) <- "RNA"
features <- c("Stfa2l1","Ptprc","Itgam","Ly6g","S100a8","Ms4a3","Mki67","Stmn1", "Elane","Mpo","Ccr2","Cd68","S100a4","Fn1","Ctss","Fcer1a","Mcpt8","Gata2","Cd69","Ms4a2","Pth1r","Twist1","Dcn","Bglap","Col1a2")
DotPlot(Axin2.combined, features = features) + RotatedAxis()+  
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=16))
```

# Figure S3: Violin plots of craniosynostosis genes in different cell types of Axin2+ mouse suture mesenchyme cells. 
```{r}
## remove the x-axis text and tick
## plot.margin to adjust the white space between each plot.
## ... pass any arguments to VlnPlot in Seurat
modify_vlnplot<- function(obj, 
                          feature, 
                          pt.size = 0, 
                          plot.margin = unit(c(0, 0, 0, 0), "cm"),
                          ...) {
  p<- VlnPlot(obj, features = feature, pt.size = pt.size, ... )  + 
    xlab("") + ylab(feature) + ggtitle("") + 
    theme(legend.position = "none",
          plot.title = element_blank(),
          axis.text.x = element_blank(), 
          axis.ticks.x = element_blank(), 
          axis.title.x = element_blank(),
          axis.title.y = element_text(size = rel(1), angle = 0), 
          axis.text.y = element_text(size = rel(1)),
          plot.margin = plot.margin ) 
  return(p)
}

## extract the max value of the y axis
extract_max<- function(p){
  ymax<- max(ggplot_build(p)$layout$panel_scales_y[[1]]$range$range)
  return(ceiling(ymax))
}


## main function
StackedVlnPlot<- function(obj, features,
                          pt.size = 0, 
                          plot.margin = unit(c(-0.75, 0, -0.75, 0), "cm"),
                          ...) {
  
  plot_list<- purrr::map(features, function(x) modify_vlnplot(obj = obj,feature = x, ...))
  
  # Add back x-axis title to bottom plot. patchwork is going to support this?
  plot_list[[length(plot_list)]]<- plot_list[[length(plot_list)]] +
    theme(axis.text.x=element_text(), axis.ticks.x = element_line())
  
  # change the y-axis tick to only max value 
  ymaxs<- purrr::map_dbl(plot_list, extract_max)
  plot_list<- purrr::map2(plot_list, ymaxs, function(x,y) x + 
                            scale_y_continuous(breaks = c(y)) + 
                            expand_limits(y = y))
  
  p<- patchwork::wrap_plots(plotlist = plot_list, ncol = 1)
  return(p)
}

DefaultAssay(Axin2.combined) <- "RNA"

features1 <- c("Ptprd","Fgfr1","Nfix","Runx2","Msx1","Tcf12","Sec24d","Ahdc1","Twist1","Fgfr2","Zic1","Il6st","Nfia","Crtap","Tgfbr2")
features2 <- c("Six1","Fto","Smo","Alpl","Efnb1","Spry1","Ift43","Kat6b","Bmp2","Dvl3","Jag1","Rasal2","Snai1","Smad6","Alx4")
features3 <- c("Bmper","Abcc9","Cyp26b1","Fgf9","Sox11","Fgfr3","Masp1","Ntrk2","Frem1","Six2","Spry4","Efna4","Wdr19")
features4 <- c("Ifrd1","Flna","Psmc2","Hnrnpk","B3gat3","Igf1r","Braf","Psmc5","Rspry1","Ptpn11","Foxo1","Kat6a")
features5 <- c("P4hb","Kansl1","Ppp1cb","Smc1a","Med13l","Slc25a24","Shoc2","Huwe1","Notch2","Axin1","Cdc45","Notch1","Arap3")
features6 <- c("Msx2","Adamtsl4","Dph1","Glis3","Ift140","Chst3","Bbs9","Rab23","Acvrl1","Acvr2a","Smurf1","Adck1","Shc4","Nphp4","Cd96","Erf","Sox6")
StackedVlnPlot(obj = Axin2.combined, features = features1)
StackedVlnPlot(obj = Axin2.combined, features = features2)
StackedVlnPlot(obj = Axin2.combined, features = features3)
StackedVlnPlot(obj = Axin2.combined, features = features4)
StackedVlnPlot(obj = Axin2.combined, features = features5)
StackedVlnPlot(obj = Axin2.combined, features = features6)
```

# Fig.2. Characterization of SuSC population by reclustering analysis of skeletogenic cell lineage in Axin2-expressing mesenchymal cells.  
```{r}
# Fig.2A and 2E
Axin2.sub<-subset(x = Axin2.combined, idents = c("NEUT1","NEUT2","MONO","MP","MC"), invert = TRUE)
DefaultAssay(Axin2.sub)<-"integrated"
Axin2.sub <- ScaleData(Axin2.sub, verbose = FALSE)
Axin2.sub <- RunPCA(Axin2.sub, npcs = 30, verbose = FALSE)
Axin2.sub <- RunUMAP(Axin2.sub, reduction = "pca", dims = 1:20)
Axin2.sub <- FindNeighbors(Axin2.sub, reduction = "pca", dims = 1:20)
Axin2.sub <- FindClusters(Axin2.sub, resolution = 0.4)
DimPlot(Axin2.sub,reduction="umap",split.by = "orig.ident", label.size=6, pt.size = 1.0)
new.cluster.ids<- c("SuSC","OB-I","OB-II","OP") 
names(x=new.cluster.ids) <-levels(x=Axin2.sub)
Axin2.sub<-RenameIdents(object = Axin2.sub,new.cluster.ids)
DimPlot(object = Axin2.sub, reduction = 'umap',label = TRUE, pt.size = 1, label.size = 6)+ NoLegend()+FontSize(x.text = 15,y.text = 15,x.title=20,y.title = 20)

# Fig.2F
DefaultAssay(Axin2.sub) <- "RNA"
FeaturePlot(Axin2.sub, feature=c("Bglap","Ibsp","Sp7"),reduction = "umap", pt.size = 0.0001, max.cutoff = 5,min.cutoff = 0, cols = c("#FFEFD5","#E6E6FA","#87CEFA","#6495ED","#4169E1","#0000CD","#000080"), label=FALSE,ncol=3)+  
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=16))
FeaturePlot(Axin2.sub, feature=c("Bglap","Ibsp","Sp7","Spp1"),reduction = "umap", pt.size = 0.0001, max.cutoff = 5,min.cutoff = 0, cols = pal, label=FALSE)+  
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=16))


# Fig.2G
featuresCellSurfaceSuSC <- c("Ptn","Tnn","Apoe","Gpc3","Fzd1","Sulf2","Egfr","Thy1","Epha4","Cdon","Ramp2","Lpl","Icam1","Thbd","Lrig3","Itga8","Nid2","Bace2","Pxdn","Scube2","Sema6c")
StackedVlnPlot(obj = Axin2.sub, features = CellSurfaceSuSC) + 
  theme(axis.text=element_text(size=18),axis.title=element_text(size=20))
```

# Fig.3. LRIG3 enrichment in the high-definition profile of SuSCs.   
```{r}
# Fig.3A
VlnPlot(object = Axin2.sub, features = c("Lrig3"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Msx1"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Msx2"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Bmper"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Bmp4"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Acvrl1"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Rab23"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Thy1"),   ncol=1, log = TRUE, pt.size = 0.5)

# Fig.3B
# Modified from Dr. Amy Merrill's code, published on Bobzin et al. 2025 Jan 15;152(2):dev204264. doi: 10.1242/dev.204264. 
FeaturePlot(Axin2.sub, features = c("Lrig3", "Cd200"), order = T, reduction = 'umap', pt.size = 0.5,  min.cutoff = 0, max.cutoff = 'q85', blend.threshold = 0, blend = T, cols = c("darkblue", "green", "magenta")) & NoAxes() &DarkTheme()
FeaturePlot(Axin2.sub, features = c("Lrig3", "Gli1"), order = T, reduction = 'umap', pt.size = 0.5,  min.cutoff = 0, max.cutoff = 'q85', blend.threshold = 0, blend = T, cols = c("darkblue", "green", "magenta")) & NoAxes() &DarkTheme()
FeaturePlot(Axin2.sub, features = c("Lrig3", "Grem1"), order = T, reduction = 'umap', pt.size = 0.5,  min.cutoff = 0, max.cutoff = 'q85', blend.threshold = 0, blend = T, cols = c("darkblue", "green", "magenta")) & NoAxes() &DarkTheme()
FeaturePlot(Axin2.sub, features = c("Lrig3", "Eng"), order = T, reduction = 'umap', pt.size = 0.5,  min.cutoff = 0, max.cutoff = 'q85', blend.threshold = 0, blend = T, cols = c("darkblue", "green", "magenta")) & NoAxes() &DarkTheme()
FeaturePlot(Axin2.sub, features = c("Lrig3", "Pdgfra"), order = T, reduction = 'umap', pt.size = 0.5,  min.cutoff = 0, max.cutoff = 'q85', blend.threshold = 0, blend = T, cols = c("darkblue", "green", "magenta")) & NoAxes() &DarkTheme()
FeaturePlot(Axin2.sub, features = c("Lrig3", "Pdgfrb"), order = T, reduction = 'umap', pt.size = 0.5,  min.cutoff = 0, max.cutoff = 'q85', blend.threshold = 0, blend = T, cols = c("darkblue", "green", "magenta")) & NoAxes() &DarkTheme()
FeaturePlot(Axin2.sub, features = c("Lrig3", "Ddr2"), order = T, reduction = 'umap', pt.size = 0.5,  min.cutoff = 0, max.cutoff = 'q85', blend.threshold = 0, blend = T, cols = c("darkblue", "green", "magenta")) & NoAxes() &DarkTheme()
FeaturePlot(Axin2.sub, features = c("Lrig3", "Thy1"), order = T, reduction = 'umap', pt.size = 0.5,  min.cutoff = 0, max.cutoff = 'q85', blend.threshold = 0, blend = T, cols = c("darkblue", "green", "magenta")) & NoAxes() &DarkTheme()
```


# Fig.S5. Reclustering of skeletogenic cell lineage in Axin2-expressing suture cells
```{r}
# Fig.S5A
VlnPlot(object = Axin2.sub, features = c("Msx1"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Thy1"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Acvrl1"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Bmper"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Rab23"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Lrig3"),   ncol=1, log = TRUE, pt.size = 0.5)

# Fig.S5B
VlnPlot(object = Axin2.sub, features = c("Runx2"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Col1a1"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Col1a2"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Alpl"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Bglap"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Sparc"),   ncol=1, log = TRUE, pt.size = 0.5)

# Fig.S5C
VlnPlot(object = Axin2.sub, features = c("Sp7"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Ibsp"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Mmp13"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Spp1"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Wnt4"),   ncol=1, log = TRUE, pt.size = 0.5)

# Fig.S5D
VlnPlot(object = Axin2.sub, features = c("Ptprc"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Mmp9"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Mpo"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Elane"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Cd14"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Cybb"),   ncol=1, log = TRUE, pt.size = 0.5)

```


# Fig.S6. Enriched Lrig3 expression in the suture mesenchyme
```{r}
# Figure S6A
VlnPlot(object = Axin2.combined, features = c("Lrig3"),   ncol=1, log = TRUE, pt.size = 0)

# Figure S6B
# Calculate percentage of gene expression in each cluster
PrctCellExpringGene <- function(object, genes, group.by = "all", assay = "RNA", datatype = "counts", threshold = 0){
    if(group.by == "all"){
        prct = unlist(lapply(genes,calc_helper, object=object, assay = assay, datatype=datatype, threshold=threshold))
        result = data.frame(Markers = genes, Cell_proportion = prct)
        return(result)
    }

    else{        
        list = SplitObject(object, group.by)
        factors = names(list)
        results = lapply(list, PrctCellExpringGene, genes=genes, assay = assay, datatype=datatype, threshold=threshold)
        results %>% reduce(full_join, by="Markers") %>% select(any_of("Markers")) -> genelist
        results %>% reduce(full_join, by="Markers") %>% select(!any_of("Markers")) %>% "colnames<-"(names(results)) -> percentages
        combined <- cbind(genelist,percentages)
        return(combined)
    }
}

calc_helper <- function(object,genes,assay,datatype,threshold){
    counts = slot(object[[assay]],datatype)
    ncells = ncol(counts)
    if(genes %in% row.names(counts)){
    round((sum(counts[genes,]>threshold)/ncells)*100,1)
    }else{return(NA)}
}

PrctCellExpringGene(Axin2.sub, genes = c("Bmpr1a","Cd200","Eng","Itgav","Thy1","Pdgfra","Pdgfrb",
                                              "Atxn1","Gli1","Grem1","Lepr","Ddr2"), 
                    group.by = "CellType", assay = "RNA", datatype = "counts", threshold = 0)
```

# Fig.S7. Distribution of stem cell gene expression in the subclusters of SC.  
```{r}
VlnPlot(object = Axin2.sub, features = c("Cd200"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Eng"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Itgav"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Pdgfra"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Pdgfrb"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Atxn1"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Gli1"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Grem1"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Lepr"),   ncol=1, log = TRUE, pt.size = 0.5)
VlnPlot(object = Axin2.sub, features = c("Ddr2"),   ncol=1, log = TRUE, pt.size = 0.5)
```

